\subsection{Pills While Blindfolded}

This a nice and accessible problem that requires some out of the box thinking. Good problem for when walking as paper would not help much. I give this problem a 1/10 in difficulty.

An evil logician has poisoned you, but gives you a way to survive. They offer you two red pills and two blue pills and to survive you need to take exactly one red and one blue pill. The only complication is that you are blindfolded. What is your strategy to survive? Standard evil logician rules apply.

\textbf{Hints:}

\begin{enumerate}
	\item There is no way to gain any information about what colour the pills are. This means your strategy would have to work for any permutation of the pills.
	\item Taking all the pills would give you exactly double the dose.
	\item Consider the same problem but with one red pill and one blue pill and needing to take half of each. Apply this to the problem with four pills.
\end{enumerate}

\textbf{Solution:}

Arrange all the pills in a line, such as ``O O O O". Instead of splitting them vertically down the middle like ``O O $|$ O O", cut them horizontally. Each half will contain exactly two lots of half a red pill and two lots of half a blue pill. This is the same amount as one red pill and one blue pill.

\textbf{Extensions and Comments:}

It is not stated whether you can or cannot cut the pills in the problems statement, but I think if you gave this information it would completely give away the solution. I think it is better to leave this as the creative option. This problem works with any number of pills and any number of colour of pills.